\section{Row-Reduced Echelon Matrices}

\Exercise1 Find all solutions to the following system of equations by
row-reducing the coefficient matrix:
\begin{alignat*}{4}
  \tfrac13x_1 &{}+{}& 2x_2 &{}-{}& 6x_3 &{}={}& 0 \\
  -4x_1 && &{}+{}& 5x_3 &{}={}& 0 \\
  -3x_1 &{}+{}& 6x_2 &{}-{}& 13x_3 &{}={}& 0 \\
  -\tfrac73x_1 &{}+{}& 2x_2 &{}-{}& \tfrac83x_3 &{}={}& 0
\end{alignat*}
\begin{solution}
  The coefficient matrix reduces as follows:
  \begin{gather*}
    \begin{bmatrix}
      \frac13 & 2 & -6 \\[3pt]
      -4 & 0 & 5 \\[3pt]
      -3 & 6 & -13 \\[3pt]
      -\frac73 & 2 & -\frac83
    \end{bmatrix}
    \xrightarrow{(1)}
    \begin{bmatrix}
      1 & 6 & -18 \\[3pt]
      -4 & 0 & 5 \\[3pt]
      -3 & 6 & -13 \\[3pt]
      -\frac73 & 2 & -\frac83
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & 6 & -18 \\[3pt]
      0 & 24 & -67 \\[3pt]
      0 & 24 & -67 \\[3pt]
      0 & 16 & -\frac{134}3
    \end{bmatrix}
    \xrightarrow{(1)} \\
    \begin{bmatrix}
      1 & 6 & -18 \\[3pt]
      0 & 1 & -\frac{67}{24} \\[3pt]
      0 & 1 & -\frac{67}{24} \\[3pt]
      0 & 16 & -\frac{134}3
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & 0 & -\frac54 \\[3pt]
      0 & 1 & -\frac{67}{24} \\[3pt]
      0 & 0 & 0 \\[3pt]
      0 & 0 & 0
    \end{bmatrix}.
  \end{gather*}
  Setting $x_3 = 24t$, we see that all solutions have the form
  \begin{equation*}
    x_1 = 30t,
    \quad
    x_2 = 67t,
    \quad\text{and}\quad
    x_3 = 24t,
  \end{equation*}
  where $t$ is an arbitrary scalar.
\end{solution}

\Exercise2 Find a row-reduced echelon matrix which is row-equivalent
to
\begin{equation*}
  A =
  \begin{bmatrix}
    1 & -i \\
    2 & 2 \\
    i & 1 + i
  \end{bmatrix}.
\end{equation*}
What are the solutions of $AX = 0$?
\begin{solution}
  Performing row-reduction on $A$ gives
  \begin{gather*}
    \begin{bmatrix}
      1 & -i \\
      2 & 2 \\
      i & 1 + i
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & -i \\
      0 & 2 + 2i \\
      0 & i
    \end{bmatrix}
    \xrightarrow{(1)}
    \begin{bmatrix}
      1 & -i \\
      0 & 1 \\
      0 & i
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & 0 \\
      0 & 1 \\
      0 & 0
    \end{bmatrix},
  \end{gather*}
  and this last matrix is in row-reduced echelon form. Therefore the
  homogeneous system $AX = 0$ has only the trivial solution
  $x_1 = x_2 = 0$.
\end{solution}

\Exercise3 Describe explicitly all $2\times2$ row-reduced echelon
matrices.
\begin{solution}
  If a $2\times2$ matrix has no nonzero rows, then it is the zero matrix,
  \begin{equation*}
    0 =
    \begin{bmatrix}
      0 & 0 \\
      0 & 0
    \end{bmatrix},
  \end{equation*}
  which is in row-reduced echelon form.

  Next, if a $2\times2$ matrix has exactly one nonzero row, then in
  order to be in row-reduced echelon form, the nonzero row must be in
  row $1$ and it must start with an entry of $1$. There are two
  possibilities,
  \begin{equation*}
    \begin{bmatrix}
      1 & a \\
      0 & 0
    \end{bmatrix}
    \quad\text{or}\quad
    \begin{bmatrix}
      0 & 1 \\
      0 & 0
    \end{bmatrix},
  \end{equation*}
  where $a$ is an arbitrary scalar.

  Lastly, if a $2\times2$ matrix in row-reduced echelon form has two
  nonzero rows, then the diagonal entries must be $1$ and the other
  entries $0$, so we get the identity matrix
  \begin{equation*}
    I =
    \begin{bmatrix}
      1 & 0 \\
      0 & 1
    \end{bmatrix}.
  \end{equation*}
  These are the only possibilities.
\end{solution}

\Exercise4 Consider the system of equations
\begin{alignat*}{4}
  x_1 &{}-{}& x_2 &{}+{}& 2x_3 &{}={}& 1 \\
  2x_1 && &{}+{}& 2x_3 &{}={}& 1 \\
  x_1 &{}-{}& 3x_2 &{}+{}& 4x_3 &{}={}& 2\rlap.
\end{alignat*}
Does this system have a solution? If so, describe explicitly all
solutions.
\begin{solution}
  We perform row-reduction on the augmented matrix:
  \begin{gather*}
    \begin{bmatrix}
      1 & -1 & 2 & 1 \\
      2 & 0 & 2 & 1 \\
      1 & -3 & 4 & 2
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & -1 & 2 & 1 \\
      0 & 2 & -2 & -1 \\
      0 & -2 & 2 & 1
    \end{bmatrix}
    \xrightarrow{(1)} \\
    \begin{bmatrix}
      1 & -1 & 2 & 1 \\
      0 & 1 & -1 & -\frac12 \\
      0 & -2 & 2 & 1
    \end{bmatrix}
    \xrightarrow{(2)}
    \begin{bmatrix}
      1 & 0 & 1 & \frac12 \\[3pt]
      0 & 1 & -1 & -\frac12 \\[3pt]
      0 & 0 & 0 & 0
    \end{bmatrix}.
  \end{gather*}
  From this we see that the original system of equations has
  solutions. All solutions are of the form
  \begin{equation*}
    x_1 = -t + \frac12,
    \quad
    x_2 = t - \frac12,
    \quad\text{and}\quad
    x_3 = t,
  \end{equation*}
  for some scalar $t$.
\end{solution}
