\skipsection
\section{Computations Concerning Subspaces}

\Exercise1 Let $s < n$ and $A$ an $s\times n$ matrix with entries in
the field $F$. Use Theorem~4 (not its proof) to show that there is a
non-zero $X$ in $F^{n\times1}$ such that $AX = 0$.
\begin{proof}
  Let $\alpha_1,\dots,\alpha_n$ denote the columns of $A$. Then each
  $\alpha_i$ is a member of the vector space $F^s$, which has
  dimension strictly less than $n$. Therefore, by Theorem~4, the
  $\alpha_i$ are necessarily linearly dependent. Thus we can write
  \begin{equation*}
    c_1\alpha_1 + c_2\alpha_2 + \cdots + c_n\alpha_n = 0
  \end{equation*}
  for $c_1,\dots, c_n$ in $F$ not all $0$. If we let
  \begin{equation*}
    X =
    \begin{bmatrix}
      c_1 \\ c_2 \\ \vdots \\ c_n
    \end{bmatrix},
  \end{equation*}
  then $AX = 0$ as required.
\end{proof}

\Exercise2 Let
\begin{equation*}
  \alpha_1 = (1, 1, -2, 1), \quad
  \alpha_2 = (3, 0, 4, -1), \quad
  \alpha_3 = (-1, 2, 5, 2).
\end{equation*}
Let
\begin{equation*}
  \alpha = (4, -5, 9, -7), \quad
  \beta = (3, 1, -4, 4), \quad
  \gamma = (-1, 1, 0, 1).
\end{equation*}
\begin{enumerate}
\item Which of the vectors $\alpha$, $\beta$, $\gamma$ are in the
  subspace of $R^4$ spanned by the $\alpha_i$?
  \begin{solution}
    Let $A$ be the $3\times4$ matrix whose $i$th row is $\alpha_i$. By
    performing row reduction on $A$, we get
    \begin{equation*}
      \begin{bmatrix}
        1 & 1 & -2 & 1 \\
        3 & 0 & 4 & -1 \\
        -1 & 2 & 5 & 2
      \end{bmatrix}
      \rightarrow
      \begin{bmatrix}
        1 & 0 & 0 & -3/13 \\
        0 & 1 & 0 & 14/13 \\
        0 & 0 & 1 & -1/13
      \end{bmatrix}.
    \end{equation*}
    A vector $\rho$ is in the row space of $A$ if and only if
    \begin{equation*}
      \rho = c_1\alpha_1 + c_2\alpha_2 + c_3\alpha_3
      = \left(c_1, c_2, c_3, -\frac{3}{13}c_1
        + \frac{14}{13}c_2 - \frac1{13}c_3\right).
    \end{equation*}
    Checking each of $\alpha$, $\beta$, and $\gamma$, we see that only
    $\alpha$ is in the row space. So $\alpha$ is in the subspace
    spanned by the $\alpha_i$, while $\beta$ and $\gamma$ are not in
    the subspace.
  \end{solution}
\item Which of the vectors $\alpha$, $\beta$, $\gamma$ are in the
  subspace of $C^4$ spanned by the $\alpha_i$?
  \begin{solution}
    Our work above is still valid in $C^4$. $\alpha_1$, $\alpha_2$,
    and $\alpha_3$ will span a larger subspace due to the scalars
    being taken from $C$ instead of $R$, but the members of this
    subspace will still have the same form as before. Thus, of the
    three vectors $\alpha$, $\beta$, and $\gamma$, only $\alpha$ is in
    the subspace.
  \end{solution}
\item Does this suggest a theorem?
  \begin{solution}
    This suggests the following theorem: let $F$ be a subfield of the
    field $E$. Let $\alpha$ be a vector in $F^n$, and let
    $\beta_1,\dots,\beta_n$ in $F^n$ span some subspace. Then $\alpha$
    is in this subspace of $F^n$ if and only if it is in the subspace
    of $E^n$ spanned by the same vectors $\beta_i$.
  \end{solution}
\end{enumerate}

\Exercise3 Consider the vectors in $R^4$ defined by
\begin{equation*}
  \alpha_1 = (-1, 0, 1, 2), \quad
  \alpha_2 = (3, 4, -2, 5), \quad
  \alpha_3 = (1, 4, 0, 9).
\end{equation*}
Find a system of homogeneous linear equations for which the space of
solutions is exactly the subspace of $R^4$ spanned by the three given
vectors.
\begin{solution}
  Let $A$ be the $3\times4$ matrix whose $i$th row is $\alpha_i$. We
  can perform row-reduction on $A$ to get
  \begin{equation*}
    R =
    \begin{bmatrix}
      1 & 0 & -1 & -2 \\[3pt]
      0 & 1 & \frac14 & \frac{11}4 \\[3pt]
      0 & 0 & 0 & 0
    \end{bmatrix}.
  \end{equation*}
  Then a vector $\rho$ in $R^4$ is in the row space of $A$ if and only
  if it has the form
  \begin{equation*}
    \rho = \left(r_1, r_2, \frac14r_2 - r_1, \frac{11}4r_2 - 2r_1\right),
  \end{equation*}
  where $r_1$ and $r_2$ are real numbers. If we label the components
  of $\rho$ as
  \begin{equation*}
    \rho = (x_1,x_2,x_3,x_4),
  \end{equation*}
  then we get the following system of equations:
  \begin{align*}
    x_3 &= \frac14x_2 - x_1 \\
    x_4 &= \frac{11}4x_2 - 2x_1.
  \end{align*}
  Or, we can rearrange these equations to write
  \begin{alignat*}{5}
    x_1 &{}-{}& \frac14x_2 &{}+{}& x_3 && &{}={}& 0 \\
    2x_1 &{}-{}& \frac{11}4x_2 && &{}+{}& x_4 &{}={}& 0.
  \end{alignat*}
  This system of equations is homogeneous and its solution set is
  precisely the subspace spanned by $\alpha_1$, $\alpha_2$, and
  $\alpha_3$.
\end{solution}

\Exercise4 In $C^3$, let
\begin{equation*}
  \alpha_1 = (1, 0, -i), \quad
  \alpha_2 = (1 + i, 1 - i, 1), \quad
  \alpha_3 = (i, i, i).
\end{equation*}
Prove that these vectors form a basis for $C^3$. What are the
coordinates of the vector $(a,b,c)$ in this basis?
\begin{solution}
  Let
  \begin{equation*}
    A =
    \begin{bmatrix}
      1 & 0 & -i \\
      1 + i & 1 - i & 1 \\
      i & i & i
    \end{bmatrix}.
  \end{equation*}
  By performing row-reduction, one can verify that $A$ is
  row-equivalent to the identity matrix. So $A$ has rank $3$ and
  $\alpha_1$, $\alpha_2$, and $\alpha_3$ are linearly independent and
  span $C^3$, as required to be a basis.

  Let the coordinates of $(a,b,c)$ in this basis be $(x,y,z)$. This
  leads to the following system of equations.
  \begin{alignat*}{4}
    x &{}+{}& (1 + i)y &{}+{}& iz &{}={}& a \\
    && (1 - i)y &{}+{}& iz &{}={}& b \\
    -ix &{}+{}& y &{}+{}& iz &{}={}& c.
  \end{alignat*}
  With a bit of effort, one may determine this system to have the
  solution
  \begin{equation*}
    \begin{bmatrix}
      x \\ y \\ z
    \end{bmatrix}
    =
    \begin{bmatrix}
      \frac{a + b - 2c}5 + \frac{4c - 2a - 2b}5i \\[3pt]
      \frac{a + b - 2c}5 + \frac{3b - 2a - c}5i \\[3pt]
      \frac{3a - 2b - c}5 - \frac{a + b + 3c}5i
    \end{bmatrix}. \qedhere
  \end{equation*}
\end{solution}

\Exercise5 Give an explicit description for the vectors
\begin{equation*}
  \beta = (b_1, b_2, b_3, b_4, b_5)
\end{equation*}
in $R^5$ which are linear combinations of the vectors
\begin{align*}
  \alpha_1 &= (1, 0, 2, 1, -1), \quad \alpha_2 = (-1, 2, -4, 2, 0) \\
  \alpha_3 &= (2, -1, 5, 2, 1), \quad \alpha_4 = (2, 1, 3, 5, 2).
\end{align*}
\begin{solution}
  Performing row-reduction on the augmented matrix
  \begin{equation*}
    A =
    \begin{bmatrix}
      1 & -1 & 2 & 2 & b_1 \\
      0 & 2 & -1 & 1 & b_2 \\
      2 & -4 & 5 & 3 & b_3 \\
      1 & 2 & 2 & 5 & b_4 \\
      -1 & 0 & 1 & 2 & b_5
    \end{bmatrix}
  \end{equation*}
  produces
  \begin{equation*}
    R =
    \begin{bmatrix}
      1 & 0 & 0 & 0 & \frac23b_1 + \frac12b_2 - \frac16b_4 - \frac12b_5 \\[3pt]
      0 & 1 & 0 & 0 & \frac76b_4 - \frac53b_1 - \frac32b_2 - \frac12b_5 \\[3pt]
      0 & 0 & 1 & 0 & \frac32b_4 - 2b_1 - \frac52b_2 - \frac12b_5 \\[3pt]
      0 & 0 & 0 & 1 & \frac43b_1 + \frac32b_2 - \frac56b_4 + \frac12b_5 \\[3pt]
      0 & 0 & 0 & 0 & b_3 + b_2 - 2b_1
    \end{bmatrix}.
  \end{equation*}
  Thus we see that $\beta = (b_1,b_2,b_3,b_4,b_5)$ is in the subspace
  spanned by $\alpha_1$, $\alpha_2$, $\alpha_3$, $\alpha_4$ if and
  only if $b_3 + b_2 - 2b_1 = 0$. For any such $\beta$, we have
  \begin{multline*}
    \beta
    = \left(\frac23b_1 + \frac12b_2 - \frac16b_4 - \frac12b_5\right)\alpha_1
    + \left(\frac76b_4 - \frac53b_1 - \frac32b_2 - \frac12b_5\right)\alpha_2 \\
    + \left(\frac32b_4 - 2b_1 - \frac52b_2 - \frac12b_5\right)\alpha_3
    + \left(\frac43b_1 + \frac32b_2 - \frac56b_4 + \frac12b_5\right)\alpha_4.
  \end{multline*}
\end{solution}

\Exercise6 Let $V$ be the real vector space spanned by the rows of the
matrix
\begin{equation*}
  A =
  \begin{bmatrix}
    3 & 21 & 0 & 9 & 0 \\
    1 & 7 & -1 & -2 & -1 \\
    2 & 14 & 0 & 6 & 1 \\
    6 & 42 & -1 & 13 & 0
  \end{bmatrix}.
\end{equation*}
\begin{enumerate}
\item Find a basis for $V$.
  \begin{solution}
    If we perform row-reduction on the matrix $A$, we get the
    row-reduced echelon matrix
    \begin{equation*}
      R =
      \begin{bmatrix}
        1 & 7 & 0 & 3 & 0 \\
        0 & 0 & 1 & 5 & 0 \\
        0 & 0 & 0 & 0 & 1 \\
        0 & 0 & 0 & 0 & 0
      \end{bmatrix}.
    \end{equation*}
    The three nonzero rows $\rho_1$, $\rho_2$, and $\rho_3$ of $R$
    form a basis for $V$.
  \end{solution}
\item Tell which vectors $(x_1, x_2, x_3, x_4, x_5)$ are elements of
  $V$.
  \begin{solution}
    If we take linear combinations of $\rho_1$, $\rho_2$, and
    $\rho_3$, we can see that the vector $(x_1,x_2,x_3,x_4,x_5)$ is in
    $V$ if and only if
    \begin{equation*}
      x_2 = 7x_1 \quad\text{and}\quad
      x_4 = 3x_1 + 5x_3. \qedhere
    \end{equation*}
  \end{solution}
\item If $(x_1,x_2,x_3,x_4,x_5)$ is in $V$ what are its coordinates in
  the basis chosen in part (a)?
  \begin{solution}
    Let $x = (x_1,x_2,x_3,x_4,x_5)$ be in $V$. If
    \begin{equation*}
      x = c_1\rho_1 + c_2\rho_2 + c_3\rho_3,
    \end{equation*}
    then we get the following system of equations:
    \begin{align*}
      c_1 &= x_1 \\
      7c_1 &= x_2 \\
      c_2 &= x_3 \\
      3c_1 + 5c_2 &= x_4 \\
      c_3 &= x_5.
    \end{align*}
    We see that $x$ has coordinates $(x_1, x_3, x_5)$ in the basis
    $\{\rho_1,\rho_2,\rho_3\}$.
  \end{solution}
\end{enumerate}
