\chapter{Preface}

This is an unofficial solution guide to the book {\em Linear Algebra},
Second Edition, by Kenneth Hoffman and Ray Kunze. It is intended for
students who are studying linear algebra using Hoffman and Kunze's
text. I encourage students who use this guide to first attempt each
exercise on their own before looking up the solution, as doing
exercises is an essential part of learning mathematics.

In writing this guide, I have avoided using techniques or results
before the point at which they are introduced in the text. My
solutions should therefore be accessible to someone who is reading
through Hoffman and Kunze for the first time.

Given the large number of exercises, errors are unavoidable in a work
such as this. I have done my best to proofread each solution, but
mistakes will get through nonetheless. If you find one, please feel
free to tell me about it via email:
\href{mailto:gkikola@gmail.com}{gkikola@gmail.com}. I appreciate any
corrections or feedback.

Please know that this guide is currently unfinished. I am slowly
working on adding the remaining chapters, but this will be done at my
own pace. If you need a solution to an exercise that I have not
included, try typing the problem statement into a web search engine
such as \href{https://www.google.com/}{Google}; it is likely that
someone else has already posted a solution.

This guide is licensed under the Creative Commons
Attribution-ShareAlike 4.0 International License. To view a copy of
this license, visit
\begin{center}
  \href{http://creativecommons.org/licenses/by-sa/4.0/}
  {http://creativecommons.org/licenses/by-sa/4.0/}
\end{center}

I am deeply grateful to the authors, Kenneth Hoffman and Ray Kunze,
for producing a well-organized and comprehensive book that is a
pleasure to read.

\begin{flushright}
  Greg Kikola\\
  \href{https://www.gregkikola.com/}{www.gregkikola.com}\\
  \href{mailto:gkikola@gmail.com}{gkikola@gmail.com}
\end{flushright}
